Outline flowCore paper

### 2008-05-14 paper/manuscript.tex
Cleaned up some of my comments.
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/flowCore@31601 bc3139a8-67e5-0310-9ffc-ced21a209358
minor fixes: typo and biblio
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/flowCore@31665 bc3139a8-67e5-0310-9ffc-ced21a209358

\section*{Introduction}
During the last several years, automation technologies have been developed
that enable the use of flow cytometry high content screening (FC-HCS)
to generate large, complex datasets in both basic and clinical research
applications \citep{references from Ryan's FC-HCS paper}. A
serious bottleneck in the intepretation of existing studies and the application of
FC-HCS to even larger, more complex problems is that data management
and data analysis methods have not advanced sufficiently far from the
methods developed for 
traditional applications of flow cytometry (FCM) to small-scale, tube-based
studies.

@@

%% EAS: I think we mention flowCore before talking about data structures and
% methods.  We should first say explicity that we have a solution for
% the missing shared platform.
The computational tools we have developed are distributed in the R 
software language (www.r-project.org) as the Bioconductor
(www.bionductor.org) package flowCore. flowCore is a freely available, highly
functional and extensible FCM data analysis platform that 
enables researchers to efficiently handle FC-HCS data and encourages open
development of tools for their coherent analysis. Among the issues
addressed by flowCore are computationally efficient data structures and
a range of specialized methods for compensation, transformation, and gating. 
flowCore runs on Windows, Mac OS X and Linux/Unix operating
systems.  In this paper, we describe the functionality and methods underlying flowCore, 
and provide examples of their use.

@@

\section*{Standard flow operations}
The basic operations in FCM analyses are typically the same: the data
need to be compensated (if not already done on the instrument side)
and transformed, and sup-populations of interest need to be selected
based on a set of (predominantly sequential) gates. All software
solutions for FCM analysis offer more or less elaborate support for
these kinds of operation, and most of them in an interactive,
GUI-driven interface. The approach we have taken in flowCore is to
abstractly describe these operations and build a set of tools to
perform them on both flow frames and flow sets. While transformation,
and to a certain extent compensation, are fairly routine operations
with only limited potential for improvement, being able to implement
new methodologies for gating of FCM data, and extending the
capabilities of flowCore through object oriented programming are
features that clearly sets our framework apart from other FCM analysis
tools. By factoring out as much of the bookkeeping as possible,
programmers can focus on the actual operations rather then having to
deal with the tedious details of data integration and
access. Third-party methods can act on their own as first-class
citizens in the analysis framework, without breaking the workflow or
the basic infrastructure. This design allows for the straightforward
extension of flowCore's capabilities, and has already fostered the
development of a number of valuable add-ons
\citep{lo2008agf,sarkar2008ufv} \textit{(plateCore reference, Perry,
Errol?)}.

@@

\secttion*{Conclusions}

>>>>

\section*{Introduction}
During the last several years, automation technologies have been developed
that enable the use of flow cytometry high content screening (FC-HCS)
to generate large, complex datasets in both basic and clinical research
applications \citep{brinkman2007hcf}. A
serious bottleneck in the intepretation of existing studies and the application of
FC-HCS to even larger, more complex problems is that data management
and data analysis methods have not advanced sufficiently far from the
methods developed for 
traditional applications of flow cytometry (FCM) to small-scale, tube-based
studies.

@@

The computational tools we have developed are distributed in the R 
software language (www.r-project.org) as the Bioconductor
(www.bionductor.org) package flowCore. flowCore is a freely available, highly
functional and extensible FCM data analysis platform that 
enables researchers to efficiently handle FC-HCS data and encourages open
development of tools for their coherent analysis. Among the issues
addressed by flowCore are computationally efficient data structures and
a range of specialized methods for compensation, transformation, and gating. 
flowCore runs on Windows, Mac OS X and Linux/Unix operating
systems.  In this paper, we describe the functionality and methods underlying flowCore, 
and provide examples of their use.

@@

\section*{Standard flow operations}
The basic operations in FCM analyses are typically the same: the data
need to be compensated (if not already done on the instrument side)
and transformed, and sup-populations of interest need to be selected
based on a set of (predominantly sequential) gates. All software
solutions for FCM analysis offer more or less elaborate support for
these kinds of operation, and most of them in an interactive,
GUI-driven interface. The approach we have taken in flowCore is to
abstractly describe these operations and build a set of tools to
perform them on both flow frames and flow sets. While transformation,
and to a certain extent compensation, are fairly routine operations
with only limited potential for improvement, being able to implement
new methodologies for gating of FCM data, and extending the
capabilities of flowCore through object oriented programming are
features that clearly sets our framework apart from other FCM analysis
tools. By factoring out as much of the bookkeeping as possible,
programmers can focus on the actual operations rather then having to
deal with the tedious details of data integration and
access. Third-party methods can act on their own as first-class
citizens in the analysis framework, without breaking the workflow or
the basic infrastructure. This design allows for the straightforward
extension of flowCore's capabilities, and has already fostered the
development of a number of valuable add-ons
\citep{lo2008agf,sarkar2008ufv} \textit{(plateCore}.

@@

\section*{Conclusions}
